Ainur:~A Framework for Repeatable End-to-End Wireless Edge Computing Testbed Research

### 2023-02-03 publications/2022Ainur/abstract.tex
fixing ainur paper

  In this paper, we present the first such mostly hardware-agnostic wireless testbed automation framework, \emph{Ainur}.
  It is designed to configure, manage, orchestrate, and deploy workloads from an end-to-end perspective.
  Ainur is built on top of cloud-native technologies such as Docker, and is provided as \gls{FOSS} to the community through the {KTH-EXPECA/Ainur} repository on GitHub.
  We demonstrate the utility of the platform with a series of scenarios, showcasing in particular its flexibility with respect to physical link definition, computation placement, and automation of arbitrarily complex experimental scenarios.
\end{abstract}

>>>>

  In this paper, we present the first such mostly hardware-agnostic wireless testbed automation framework, \emph{Ainur}.
  It is designed to configure, manage, orchestrate, and deploy workloads from an end-to-end perspective.
  Ainur is built on top of cloud-native technologies such as Docker, and is provided as \gls{FOSS} to the community through the \href{https://github.com/KTH-EXPECA/Ainur}{KTH-EXPECA/Ainur} repository on GitHub.
  We demonstrate the utility of the platform with a series of scenarios, showcasing in particular its flexibility with respect to physical link definition, computation placement, and automation of arbitrarily complex experimental scenarios.
\end{abstract}

### 2023-02-03 publications/2022Ainur/ainur.tex
fixing ainur paper

\section{The Ainur Framework}\label{sec:ainur}

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\linewidth]{publications/2022Ainur/figures/overview.jpg}
    \caption{Layered structure of an Ainur experiment}\label{fig:overview}
\end{figure}

\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{publications/2022Ainur/figures/flow2.png}
    \caption{Lifecycle of an experimental run in Ainur. Blocks in orange are optional.}\label{fig:flow}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=.8\textwidth]{publications/2022Ainur/figures/network.png}
    \caption{Network structure assumed by Ainur.}\label{fig:network}
\end{figure*}

@@

Conceptually, in Ainur, an experiment is decomposed into a layered structure, consisting of
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item the distributed application (workload)
    \item computation hosts
    \item communication networks connecting the hosts
\end{enumerate*}.
In \cref{fig:overview}, the workload is represented by a client-server process pair.
These could correspond, for instance, to the emulation of an inverted pendulum and a matching controller.
However, Ainur makes no assumptions about the nature of the workloads deployed on the framework, and virtually any process or combination of processes can be used.

Hosts correspond to either bare-metal machines or cloud instances on \gls{AWS} \gls{EC2}.
Users are free to combine and interconnect these in any configuration.
Ainur can provision wired networks over ethernet, and supports a number of software-defined wireless communication stacks, currently including WiFi, 4G \gls{LTE}, and 5G.
To realize these various wireless communication protocols, it is assumed that the testbed is equipped with \gls{USRP} \gls{SDR}s and some computation hosts dedicated to signal processing and/or radio management.

@@

    \item[Containers:]
    We leverage Docker containers to great extent for the virtualization and orchestration of workloads, as well as the encapsulation and management of complex network configurations.
    See \cref{ssec:containers} for more details.

@@

    \item[\gls{OAI}:]
    \gls{OAI} is an open-source project that implements 3GPP technology on general purpose \texttt{x86} computing hardware and off-the-shelf \glspl{SDR} like the \gls{USRP}~\cite{KALTENBERGER2020107284}.
    Ainur can deploy, configure, and manage \gls{OAI} software components that implement 4G \gls{LTE} and 5G \gls{NR}.

    \item[Mango Communications 802.11:]
    Finally, the Mango Communications project implements real-time 802.11 (WiFi) MAC and PHY in Xilinx \glspl{FPGA}.
    It can be used on a variety of hardware platforms including \gls{USRP} \glspl{SDR}, and is employed in Ainur to provision end-to-end WiFi links.
\end{description}

@@

    \item[Physical Layer:] 
    Creates and deploys the underlying physical connections of the workload data network.
    This layer interacts with hardware such as managed switches and \glspl{SDR} (and their associated computation hosts) to create links between the desired workload hosts.
    These links correspond to ethernet, WiFi, and even an entire software-defined 4G \gls{LTE} and 5G networks.

@@

\subsection{Containerization}\label{ssec:containers}

@@

\begin{figure}
    \centering
    \includegraphics[width=.8\columnwidth]{publications/2022Ainur/figures/network-minimal.png}
    \caption{Minimal testbed setup on which Ainur can be deployed.}\label{fig:network:minimal}
\end{figure}

@@

A full-featured, and rather complex deployment is depicted in \cref{fig:network}, including a number of optional testbed features such as offload to the cloud through \gls{VPN} gateways.
However, a minimal a setup can be achieved with a control host, at least two workload hosts, and a couple of managed (smart) switches, as illustrated in \cref{fig:network:minimal}.
The only requirements for the aforementioned hosts are that a general purpose Linux distribution such as Ubuntu can be deployed on them, and --- in the case of the workload hosts --- the presence of two network interfaces.
Thus, a minimal testbed can be achieved cheaply and quickly by employing small single-board computers such as Raspberry Pi, Jetson Nano, or BeagleBone units.
Finally, the configuration and deployment of an Ainur-compatible testbed can be easily automated with tools such as Ansible, and playbooks for this effect can be made available to the community upon request.

>>>>

\section{The Ainur Framework}\label{paper:olguinmunoz2022airnur:ainur}

\begin{figure}[t]
    \centering
    \includegraphics[width=\textwidth]{publications/2022Ainur/figures/overview}
    \caption{Layered structure of an Ainur experiment}\label{paper:olguinmunoz2022airnur:fig:overview}
\end{figure}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{publications/2022Ainur/figures/flow2}
    \caption{Lifecycle of an experimental run in Ainur. Blocks in orange are optional.}\label{paper:olguinmunoz2022airnur:fig:flow}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{publications/2022Ainur/figures/network}
    \caption{Network structure assumed by Ainur.}\label{paper:olguinmunoz2022airnur:fig:network}
\end{figure*}

@@

Conceptually, in Ainur, an experiment is decomposed into a layered structure, consisting of
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item the distributed application (workload)
    \item computation hosts
    \item communication networks connecting the hosts
\end{enumerate*}.
In \cref{paper:olguinmunoz2022airnur:fig:overview}, the workload is represented by a client-server process pair.
These could correspond, for instance, to the emulation of an inverted pendulum and a matching controller.
However, Ainur makes no assumptions about the nature of the workloads deployed on the framework, and virtually any process or combination of processes can be used.

Hosts correspond to either bare-metal machines or cloud instances on \gls{AWS} \gls{EC2}.
Users are free to combine and interconnect these in any configuration.
Ainur can provision wired networks over ethernet, and supports a number of software-defined wireless communication stacks, currently including Wi-Fi, 4G \gls{LTE}, and 5G.
To realize these various wireless communication protocols, it is assumed that the testbed is equipped with \gls{USRP} \gls{SDR}s and some computation hosts dedicated to signal processing and/or radio management.

@@

    \item[Containers:]
    We leverage Docker containers to great extent for the virtualization and orchestration of workloads, as well as the encapsulation and management of complex network configurations.
    See \cref{paper:olguinmunoz2022airnur:containers} for more details.

@@

    \item[\gls{OAI}:]
    \gls{OAI} is an open-source project that implements 3GPP technology on general purpose \texttt{x86} computing hardware and off-the-shelf \glspl{SDR} like the \gls{USRP}~\cite{kaltenberger2020openairinterface}.
    Ainur can deploy, configure, and manage \gls{OAI} software components that implement 4G \gls{LTE} and 5G \gls{NR}.

    \item[Mango Communications 802.11:]
    Finally, the Mango Communications project implements real-time 802.11 (Wi-Fi) MAC and PHY in Xilinx\\\glspl{FPGA}.
    It can be used on a variety of hardware platforms including \gls{USRP} \glspl{SDR}, and is employed in Ainur to provision end-to-end Wi-Fi links.
\end{description}

@@

    \item[Physical Layer:] 
    Creates and deploys the underlying physical connections of the workload data network.
    This layer interacts with hardware such as managed switches and \glspl{SDR} (and their associated computation hosts) to create links between the desired workload hosts.
    These links correspond to ethernet, Wi-Fi, and even an entire software-defined 4G \gls{LTE} and 5G networks.

@@

\subsection{Containerization}\label{paper:olguinmunoz2022airnur:containers}

@@

\begin{figure}
    \centering
    \includegraphics[height=25em]{publications/2022Ainur/figures/network-minimal}
    \caption{Minimal testbed setup on which Ainur can be deployed.}\label{paper:olguinmunoz2022airnur:fig:network:minimal}
\end{figure}

@@

A full-featured, and rather complex deployment is depicted in \cref{paper:olguinmunoz2022airnur:fig:network}, including a number of optional testbed features such as offload to the cloud through \gls{VPN} gateways.
However, a minimal a setup can be achieved with a control host, at least two workload hosts, and a couple of managed (smart) switches, as illustrated in \cref{paper:olguinmunoz2022airnur:fig:network:minimal}.
The only requirements for the aforementioned hosts are that a general purpose Linux distribution such as Ubuntu can be deployed on them, and --- in the case of the workload hosts --- the presence of two network interfaces.
Thus, a minimal testbed can be achieved cheaply and quickly by employing small single-board computers such as Raspberry Pi, Jetson Nano, or BeagleBone units.
Finally, the configuration and deployment of an Ainur-compatible testbed can be easily automated with tools such as Ansible, and playbooks for this effect can be made available to the community upon request.

### 2023-02-03 publications/2022Ainur/demo.tex
fixing ainur paper

\begin{figure*}
    \centering
    \includegraphics[width=.8\textwidth]{publications/2022Ainur/figures/demo_hardware.png}
    \caption[caption]{%
        The possible workload offloading setups and physical layer configurations used in the demo.
        Not pictured, in Phy Option 2:
        \begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
            \item \gls{SDR} processing hosts
            \item additional base-station host for the Core Network and \gls{eNodeB}
        \end{enumerate*}.
    }\label{fig:democonfigs}
\end{figure*}

\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[height=8em]{publications/2022Ainur/figures/demo_workload_1.png}
        \caption{Workload 1: \gls{MNIST} image classifier}\label{fig:wkld:mnist}
    \end{subfigure}%
    \vspace{1em}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[height=8em]{publications/2022Ainur/figures/demo_workload_2.png}
        \caption{Workload 2: inverted pendulum \gls{NCS}}\label{fig:wkld:ncs}
    \end{subfigure}
    \caption{Demonstration workloads}\label{fig:wkld}
\end{figure}

\section{Demo}\label{sec:demo}

@@

\Cref{fig:democonfigs} illustrates the possible configurations for the testbed.
Workloads are deployed on client hosts, and computation is offloaded either to an edge server (cloudlet), or to \gls{AWS} \gls{EC2} instances on the cloud.
Communication between the client- and server-sides of the workload will occur over one of three possible physical link layer setups:
\begin{description}
    \item[WiFi:] \pgls{SDR} is configured as an IEEE 802.11n access point, and client hosts connect to it using on-board WiFi.
    \item[4G \gls{LTE}:] \pgls{SDR} is configured as an 4G \gls{LTE} base station, and another is configured as an 4G \gls{LTE} \gls{UE}.
    Together these radios act as an \gls{LTE} bridge between the client-side and the server-side of the network, and all client-server traffic is routed through them.
    \item[Ethernet:] connects everything through plain ethernet.
\end{description}

The number of workload instances (and therefore client hosts) deployed in each execution of this demonstration will range from \numrange[]{1}{10}.
Workloads deployed to the cloud will be able to target any of \gls{AWS}'s datacenters.

We will employ two different workloads for this demonstration, illustrated in \cref{fig:wkld}.
The first of these (\cref{fig:wkld:mnist}) consists of a proof-of-concept web application which implements a simple classifier to identify hand-drawn digits from the widely-used \gls{MNIST} dataset~\cite{mnist}.
The application consists of a client with a web interface, and \pgls{HTTP} server which responds to requests from the client and performs the actual image recognition.
This workload is intended to showcase in an interactive manner the effects of placing computation at different points of the network, and how Ainur simplifies these deployments.

The second workload (\cref{fig:wkld:ncs}) corresponds to \pgls{NCS} balancing an inverted pendulum, implemented on a software framework for the emulation of \glspl{NCS} using cloud-native technologies~\cite{cleave}.
It consists of an emulation of the physical inverted pendulum system plant and a software-implemented proportional-differential controller.
These components communicate with each other over the \gls{UDP}.
This workload will be used to showcase the utility of Ainur for automating the execution of batches of experiments potentially including multiple different clients, servers, and physical layers.

@@

This demonstration will be performed on a testbed consisting of
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and finally }}]
    \item \num{10} Raspberry Pi 4 Model B boards, acting as client-side workload hosts
    \item a Raspberry Pi 4 Model B acting as \gls{VPN} gateway for the workload network
    \item a Raspberry Pi 4 Model B acting as \gls{VPN} gateway for the management network, as well as hosting the necessary \gls{NTP} and \gls{DNS} server software
    \item a \texttt{i386} workstation, acting as an edge-side workload host
    \item a configurable number of  \gls{AWS} \gls{EC2} cloud instances acting as cloud servers
    \item a separate \texttt{i386} workstation hosting the Fluent server and on which Ainur is deployed as well.
\end{enumerate*}
These nodes are interconnected using a combination of managed switches, \glspl{SDR}, and \gls{VPN} gateways; please refer to \cref{fig:network} for an architectural overview of this setup.

@@

\begin{figure}
    \centering
    \includegraphics[width=.65\columnwidth]{publications/2022Ainur/figures/demo_mockup.png}
    \caption{Screenshot of the user interface of demo workload 1. After deciding on the placement of the compute backend and the type of physical layer connecting client and backend, participants will use this interface to interact with the system.}\label{fig:mockup}
\end{figure}

Participants will be asked to choose
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item a location to offload computation to (local (i.e.\ no offloading), edge, or any \gls{AWS} datacenter)
    \item a physical layer for the first hop of the network (WiFi, 4G \gls{LTE}, or ethernet).
\end{enumerate*}
Through the Ainur command line, we will deploy a single client-server pair according to the specifications.
Next, participants will be able to access the client web interface and interact with the server by requesting classification of images from the \gls{MNIST} database.
The server will respond to these, and the assigned labels will be visible on the web interface together with timing statistics for each request, see \cref{fig:mockup}.

>>>>

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{publications/2022Ainur/figures/demo_hardware}
    \caption[caption]{%
        The possible workload offloading setups and physical layer configurations used in the demo.
        Not pictured, in Phy Option 2:
        \begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
            \item \gls{SDR} processing hosts
            \item additional base-station host for the Core Network and \gls{eNodeB}
        \end{enumerate*}.
    }\label{paper:olguinmunoz2022airnur:fig:democonfigs}
\end{figure*}

\begin{figure}
    \centering
    \begin{subfigure}{.45\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{publications/2022Ainur/figures/demo_workload_1}
        \caption{Workload 1: \acs{MNIST} image classifier}\label{paper:olguinmunoz2022airnur:fig:wkld:mnist}
    \end{subfigure}%
    \hfill%
%    \vspace{1em}
    \begin{subfigure}{.45\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{publications/2022Ainur/figures/demo_workload_2}
        \caption{Workload 2: inverted pendulum \acs{NCS}}\label{paper:olguinmunoz2022airnur:fig:wkld:ncs}
    \end{subfigure}
    \caption{Demonstration workloads}\label{paper:olguinmunoz2022airnur:fig:wkld}
\end{figure}

\section{Demo}\label{paper:olguinmunoz2022airnur:demo}

@@

\Cref{paper:olguinmunoz2022airnur:fig:democonfigs} illustrates the possible configurations for the testbed.
Workloads are deployed on client hosts, and computation is offloaded either to an edge server (cloudlet), or to \gls{AWS} \gls{EC2} instances on the cloud.
Communication between the client- and server-sides of the workload will occur over one of three possible physical link layer setups:
\begin{description}
    \item[WiFi:] \pgls{SDR} is configured as an IEEE 802.11n access point, and client hosts connect to it using on-board Wi-Fi.
    \item[4G \gls{LTE}:] \pgls{SDR} is configured as an 4G \gls{LTE} base station, and another is configured as an 4G \gls{LTE} \gls{UE}.
    Together these radios act as an \gls{LTE} bridge between the client-side and the server-side of the network, and all client-server traffic is routed through them.
    \item[Ethernet:] connects everything through plain ethernet.
\end{description}

The number of workload instances (and therefore client hosts) deployed in each execution of this demonstration will range from \numrange[]{1}{10}.
Workloads deployed to the cloud will be able to target any of \gls{AWS}s datacenters.

We will employ two different workloads for this demonstration, illustrated in \cref{paper:olguinmunoz2022airnur:fig:wkld}.
The first of these (\cref{paper:olguinmunoz2022airnur:fig:wkld:mnist}) consists of a proof-of-concept web application which implements a simple classifier to identify hand-drawn digits from the widely-used \gls{MNIST} dataset~\cite{deng2012mnist}.
The application consists of a client with a web interface, and \pgls{HTTP} server which responds to requests from the client and performs the actual image recognition.
This workload is intended to showcase in an interactive manner the effects of placing computation at different points of the network, and how Ainur simplifies these deployments.

The second workload (\cref{paper:olguinmunoz2022airnur:fig:wkld:ncs}) corresponds to \pgls{NCS} balancing an inverted pendulum, implemented on a software framework for the emulation of \glspl{NCS} using cloud-native technologies~\cite{olguinmunoz2022cleave}.
It consists of an emulation of the physical inverted pendulum system plant and a software-implemented proportional-differential controller.
These components communicate with each other over the \gls{UDP}.
This workload will be used to showcase the utility of Ainur for automating the execution of batches of experiments potentially including multiple different clients, servers, and physical layers.

@@

This demonstration will be performed on a testbed consisting of
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and finally }}]
    \item \num{10} Raspberry Pi 4 Model B boards, acting as client-side workload hosts
    \item a Raspberry Pi 4 Model B acting as \gls{VPN} gateway for the workload network
    \item a Raspberry Pi 4 Model B acting as \gls{VPN} gateway for the management network, as well as hosting the necessary \gls{NTP} and \gls{DNS} server software
    \item a \texttt{i386} workstation, acting as an edge-side workload host
    \item a configurable number of  \gls{AWS} \gls{EC2} cloud instances acting as cloud servers
    \item a separate \texttt{i386} workstation hosting the Fluent server and on which Ainur is deployed as well.
\end{enumerate*}
These nodes are interconnected using a combination of managed switches, \glspl{SDR}, and \gls{VPN} gateways; please refer to \cref{paper:olguinmunoz2022airnur:fig:network} for an architectural overview of this setup.

@@

\begin{figure}
    \centering
    \includegraphics[height=25em]{publications/2022Ainur/figures/demo_mockup}
    \caption{Screenshot of the user interface of demo workload 1. After deciding on the placement of the compute backend and the type of physical layer connecting client and backend, participants will use this interface to interact with the system.}\label{paper:olguinmunoz2022airnur:fig:mockup}
\end{figure}

Participants will be asked to choose
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item a location to offload computation to (local (i.e.\ no offloading), edge, or any \gls{AWS} datacenter)
    \item a physical layer for the first hop of the network (Wi-Fi, 4G \gls{LTE}, or ethernet).
\end{enumerate*}
Through the Ainur command line, we will deploy a single client-server pair according to the specifications.
Next, participants will be able to access the client web interface and interact with the server by requesting classification of images from the \gls{MNIST} database.
The server will respond to these, and the assigned labels will be visible on the web interface together with timing statistics for each request, see \cref{paper:olguinmunoz2022airnur:fig:mockup}.

### 2023-02-03 publications/2022Ainur/introduction.tex
fixing ainur paper

Over the last years, several small- to mid-scale testbeds have emerged that leverage a large degree of freedom with respect to hardware and software, for example, the 
\begin{enumerate*}[itemjoin={{, }}, itemjoin*={{, and }}]
    \item \gls{COSMOS}
    \item \gls{POWDER}
    \item \gls{ARA}
    \item Drexel Grid \gls{SDR}
\end{enumerate*} testbeds.
\gls{COSMOS} is a testbed spanning an area of roughly \num{1} square mile (\SI{2.6}{\kilo\meter\squared}) featuring \glspl{SDR}, \si{\milli\meter}-wave equipment, optical fibers, cloud integration, and compute for core network functionality and application data processing~\cite{Cosmos1,cosmos2}.
It contains over \num{200} rooftop, intermediate, and mobile nodes, and is controlled and managed by a central node.
\gls{COSMOS} relies on the \gls{OMF} (originally developed for ORBIT~\cite{orbit}), and employs the \gls{OEDL}, a domain-specific imperative language based on Ruby, for experiment development and definition.

\gls{POWDER} promises research on wireless and mobile networks with a level of programmability down to the waveform~\cite{powder}.
The testbed spans a \SI{15}{\kilo\meter\squared} area and features about \num{15} fixed programmable radio nodes, based on off-the-shelves \glspl{SDR} and featuring edge-like compute capabilities and integration with cloud resources, which interact with \num{50} mobile nodes.
\gls{POWDER} experiments are defined and developed in \emph{profiles}, which correspond to \gls{VM} images containing the necessary software and configurations.
These profiles are defined through using \gls{RSpec}\footnote{\url{https://groups.geni.net/geni/wiki/GENIExperimenter/RSpecs}} documents.

@@

Finally, the \emph{Drexel Grid \gls{SDR} Testbed} features \glspl{SDR} that connect either over-the-air, through a channel emulator, or over a combination of the two, to facilitate realistic and reproducible experimentation~\cite{DrexelGrid}.
Primarily intended for \gls{SDR}-centric research, it does not integrate any core, cloud or edge components.
However, the testbed extensively employs the \gls{LXC} runtime for the deployment of both experimental code and \gls{SDR} software, which affords users great freedom when it comes to the development of experiments.

@@

To the best of our knowledge, CloudRAFT~\cite{cloudraft} is the only work to tackle (to a certain extent) this challenge.
CloudRAFT corresponds to a cloud-based framework for mobile network experimentation, with a focus on simplifying the management of testbed resources.
The goal of this project is to integrate, coordinate, share, and improve upon existing testbeds; and employs pre-built \glspl{VM} containing the necessary software for experiments.
Although it provides some automation for testbed resource provisioning and experiment execution, its focus is largely rather on the sharing and partitioning of testbed systems.
Testbeds currently working with CloudRAFT include a variety of domain-specific setups, including \pgls{SDR}-based testbed as a well as a ground vehicular robot for mobility-related experimentation.

@@

The rest of this paper is structured as follows.
\cref{sec:ainur} presents the framework as well as the key concepts and technologies supporting its design and architecture.
This section also discusses briefly the assumptions made about the underlying hardware on which the framework is set to run, and we present an overview of our experimental testbed.
Next, \cref{sec:demo} describes two demonstration procedures through which we will showcase the flexibility and potential of this tool for the automatic, repeatable, end-to-end experimentation in the context of wireless testbeds.
Finally, in \cref{sec:conclusion} we summarize our contributions, discuss future directions, and conclude this paper.

>>>>

Over the last years, several small- to mid-scale testbeds have emerged that leverage a large degree of freedom with respect to hardware and software, for example, the 
\begin{enumerate*}[itemjoin={{, }}, itemjoin*={{, and }}]
    \item \gls{COSMOS}
    \item \gls{POWDER}
    \item \gls{ARA}
    \item Drexel Grid \gls{SDR}
\end{enumerate*} testbeds.
\gls{COSMOS} is a testbed spanning an area of roughly \num{1} square mile (\SI{2.6}{\kilo\meter\squared}) featuring \glspl{SDR}, \si{\milli\meter}-wave equipment, optical fibers, cloud integration, and compute for core network functionality and application data processing~\cite{yu2019cosmos,raychaudhuri2020challenge}.
It contains over \num{200} rooftop, intermediate, and mobile nodes, and is controlled and managed by a central node.
\gls{COSMOS} relies on the \gls{OMF} (originally developed for ORBIT~\cite{ott2005orbit}), and employs the \gls{OEDL}, a domain-specific imperative language based on Ruby, for experiment development and definition.

\gls{POWDER} promises research on wireless and mobile networks with a level of programmability down to the waveform~\cite{breen2020powder}.
The testbed spans a \SI{15}{\kilo\meter\squared} area and features about \num{15} fixed programmable radio nodes, based on off-the-shelves \glspl{SDR} and featuring edge-like compute capabilities and integration with cloud resources, which interact with \num{50} mobile nodes.
\gls{POWDER} experiments are defined and developed in \emph{profiles}, which correspond to \gls{VM} images containing the necessary software and configurations.
These profiles are defined through using \gls{RSpec}\footnote{\url{https://groups.geni.net/geni/wiki/GENIExperimenter/RSpecs}} documents.

@@

Finally, the \emph{Drexel Grid \gls{SDR} Testbed} features \glspl{SDR} that connect either over-the-air, through a channel emulator, or over a combination of the two, to facilitate realistic and reproducible experimentation~\cite{dandekar2019grid}.
Primarily intended for \gls{SDR}-centric research, it does not integrate any core, cloud or edge components.
However, the testbed extensively employs the \gls{LXC} runtime for the deployment of both experimental code and \gls{SDR} software, which affords users great freedom when it comes to the development of experiments.

@@

To the best of our knowledge, CloudRAFT~\cite{moorthy2022cloudraft} is the only work to tackle (to a certain extent) this challenge.
CloudRAFT corresponds to a cloud-based framework for mobile network experimentation, with a focus on simplifying the management of testbed resources.
The goal of this project is to integrate, coordinate, share, and improve upon existing testbeds;
and employs pre-built \glspl{VM} containing the necessary software for experiments.
Although it provides some automation for testbed resource provisioning and experiment execution, its focus is largely rather on the sharing and partitioning of testbed systems.
Testbeds currently working with CloudRAFT include a variety of domain-specific setups, including \pgls{SDR}-based testbed as a well as a ground vehicular robot for mobility-related experimentation.

@@

The rest of this paper is structured as follows.
\cref{paper:olguinmunoz2022airnur:ainur} presents the framework as well as the key concepts and technologies supporting its design and architecture.
This section also discusses briefly the assumptions made about the underlying hardware on which the framework is set to run, and we present an overview of our experimental testbed.
Next, \cref{paper:olguinmunoz2022airnur:demo} describes two demonstration procedures through which we will showcase the flexibility and potential of this tool for the automatic, repeatable, end-to-end experimentation in the context of wireless testbeds.
Finally, in \cref{paper:olguinmunoz2022airnur:conclusion} we summarize our contributions, discuss future directions, and conclude this paper.
